\documentclass{article}

\usepackage[handout]{/Users/jared.mcbride/Library/CloudStorage/OneDrive-SouthernVirginiaUniversity/Teaching/McB/McB}

\def \meetingtimes{MWF 8:00 -- 8:50}
\def \meetingplace{100 Craton Hall}
\def \course{September 2025}
\def \semester{Fall 2025}
\def \subtitle{Jared McBride}
\def \date{\bf Curriculum Vitae}
\def \title{}

\begin{document}

\maketitle

\section*{Education}
\textbf{Ph.D. Applied Mathematics}\newline
University of Arizona, Tucson, Arizona. August 2023.
\\ Graduate minor in {\bf Systems and Industrial Engineering}. 
\\Dissertation: \emph{The Estimation of High-contrast Spectra via Iterated Whitening}. 
\\Advisor: Kevin Lin.

\textbf{M.S. Applied Mathematics}\newline
University of Arizona, Tucson, Arizona. May 2018. 
\\Advisor: Kevin Lin.

\textbf{M.S. Mathematics}\newline
Brigham Young University, Provo, Utah. August 2016. 
\\Thesis: \emph{Steady State Configurations of Cells Connected by Cadherin Sites}. 
\\Advisor: John Dallon.

\textbf{B.S. Mathematics and Mathematics Education}\newline
Brigham Young University. April 2012.

\section*{Employment}
\textbf{Southern Virginia University}, Buena Vista, Virginia\newline
\emph{Assistant Professor of Math} 2023 - present

\textbf{Mathematics Department, UA}, Tucson, Arizona\newline
\emph{Instructor} 2023

\textbf{Mathematics Department, UA}, Tucson, Arizona\newline
\emph{Graduate Teaching Assistant} 2016 - 2022

\textbf{Chemistry \& Biochemistry Department, UA}, Tucson, Arizona\newline
\emph{GRE Preparation Math Tutor} for the Maximizing Access to Research Careers (MARC) Program 2016

\textbf{Mathematics Department, BYU}, Provo, Utah\newline
\emph{Research Assistant} to Dr. J. Humphreys 2015 - 2016\newline
Proofread several chapters of \emph{Foundations of Applied Mathematics, Volume I: Mathematical Analysis}

\textbf{Mathematics Department, BYU}, Provo, Utah\newline
\emph{Graduate Teaching Assistant} 2014 - 2016

\textbf{Utah Valley University}, Orem, Utah\newline
\emph{Assistant Coordinator} in the Mathematics Tutorial Lab 2013 - 2014

\textbf{Vernal Junior High}, Vernal, Utah\newline
\emph{Math Teacher} (8th grade) 2012 - 2013

\section*{Teaching}
\subsection*{Southern Virginia University}
(August 2023 - present)
\begin{longtable}{|>{\centering\arraybackslash}m{2cm}|>{\arraybackslash}m{6cm}|>{\centering\arraybackslash}m{2cm}|}
\hline
\textbf{Semester} & \textbf{Course} & \textbf{Units} \\
\hline
\hline
\endhead
Fall 2025 & CSC 340: Artificial Intelligence & 3 \\
\hline
Fall 2025 & MAT 343: Linear Algebra & 3 \\
\hline
Fall 2025 & MAT 221: Statistics & 3  ($\times 2$) \\
\hline
Fall 2025 & MAT 115: College Algebra & 3 \\
\hline
\hline
Summer 2025 & MAT 221: Statistics & 3 \\
\hline 
Spring 2025 & CSC 333: Computational Data Science & 3 \\
\hline
Spring 2025 & MAT 498: Mathematics Capstone & 1 \\
\hline
Spring 2025 & MAT 221: Statistics & 3 ($\times 3$) \\
\hline
Fall 2024 & MAT 341: Real Analysis & 3 \\
\hline
Fall 2024 & MAT 343: Linear Algebra & 3 \\
\hline
Fall 2024 & MAT 221: Statistics & 3 ($\times 3$) \\
\hline
\hline
Spring 2024 & MAT 410: Introduction to Numerical Analysis & 3 \\
\hline
Spring 2024 & MAT 341: Calculus III & 3 \\
\hline 
Spring 2024 & MAT 221: Statistics & 3 ($\times 2$) \\
\hline
Spring 2024 & MAT 116: Trigonometry & 1 \\
\hline
Spring 2024 & MAT 115: College Algebra & 3 \\
\hline
Fall 2023 & MAT 343: Linear Algebra & 3 \\
\hline
Fall 2023 & MAT 221: Statistics & 3 ($\times 2$) \\
\hline
Fall 2023 & MAT 115: College Algebra & 3 \\
\hline
\end{longtable}

\subsection*{Department of Mathematics, University of Arizona}
(Fall 2016 - Spring 2023)
\begin{longtable}{|>{\centering\arraybackslash}m{2cm}|>{\arraybackslash}m{6cm}|>{\centering\arraybackslash}m{2cm}|}
\hline
\textbf{Semester} & \textbf{Course} & \textbf{Units} \\
\hline
\hline
\endhead
Spring 2023 & Math 116: Business Calculus & 3 ($\times 2$) \\
\hline
\hline
Spring 2022 & Math 196V: Vector Calculus Supplement & 1 \\
\hline
Fall 2021 & Math 196V: Vector Calculus Supplement & 1 \\
\hline
\hline
Spring 2021 & Math 129: Second Semester Calculus & 3 \\
\hline
Fall 2020 & Math 122B: First Semester Calculus & 4 \\
\hline
\hline
Spring 2020 & Math 122A: Functions of Calculus & 1 ($\times 2$) \\
\hline
Fall 2019 & Math 122A: Functions of Calculus & 1 ($\times 2$) \\
\hline
\hline
Spring 2019 & Math 122A: Functions of Calculus & 1 ($\times 2$) \\
\hline
Fall 2018 & Math 120R: Precalculus & 4 \\
\hline
\hline
Summer 2018 & Math 116: Business Calculus (online) & 3 \\
\hline
Spring 2018 & Math 116: Business Calculus & 3 \\
\hline
Fall 2017 & Math 120R: Precalculus & 4 \\
\hline
\hline
Spring 2017 & Math 116: Business Calculus & 3 \\
\hline
Fall 2016 & Math 112: College Algebra & 3 ($\times 2$) \\
\hline
\end{longtable}

\subsection*{Department of Mathematics, Brigham Young University}
(Spring 2014 - Summer 2015)
\begin{longtable}{|>{\centering\arraybackslash}m{2cm}|>{\arraybackslash}m{6cm}|>{\centering\arraybackslash}m{2cm}|}
\hline
\textbf{Semester} & \textbf{Course} & \textbf{Units} \\
\hline
\hline
\endhead
Fall 2015 & Math 110: College Algebra & 3 \\
\hline
\hline
Summer 2015 & Math 112: Calculus I & 4 \\
\hline
Winter 2015 & Math 102: Quantitative Reasoning & 3 \\
\hline
Fall 2014 & Math 110: College Algebra & 3 \\
\hline
\hline
Spring 2014 & Math 102: Quantitative Reasoning & 3 \\
\hline
\end{longtable}

\subsection*{Teaching Assistant}
\emph{In all courses I assisted the instructor in varying capacities ranging from grading, holding office hours, holding problem sessions, etc.}

\subsubsection*{Department of Mathematics, University of Arizona}
(Fall 2018 - Spring 2022)
\begin{longtable}{|>{\centering\arraybackslash}m{2cm}|>{\arraybackslash}m{6cm}|>{\arraybackslash}m{6cm}|}
\hline
\textbf{Semester} & \textbf{Course} & \textbf{Instructor} \\
\hline
\endhead
Spring 2022 & Math 485: Intro to Mathematical Modeling & Joceline Lega \\
\hline
Spring 2022 & Math 584B: Principles of Analysis & Shankar Venkataramani \\
\hline
Fall 2021 & Math 584A: Principles of Analysis & Shankar Venkataramani \\
\hline
\hline
Spring 2021 & Math 527B: Principles of Analysis & Shankar Venkataramani \\
\hline
Fall 2020 & Math 527A: Principles of Analysis & Shankar Venkataramani \\
\hline
\hline
Spring 2020 & Math 527B: Principles of Analysis & Shankar Venkataramani \\
\hline
Fall 2019 & Math 527A: Principles of Analysis & Shankar Venkataramani \\
\hline
\hline
Summer 2019 & Co-facilitated program sponsored qualifying exam prep course & \\
\hline
Spring 2019 & Math 527B: Principles of Analysis & Leonid Friedlander \\
\hline
\hline
Summer 2018 & Co-facilitated program sponsored qualifying exam prep course & \\
\hline
Fall 2018 & Math 527A: Principles of Analysis & Leonid Friedlander \\
\hline
\end{longtable}

\subsection*{Course Development}

\textbf{Artificial Intelligence} (\emph{Southern Virginia University CSC 340, upper division computer science course artificial intelligence}) I developed the curriculum to focus on developing deep conceptual understanding of neural networks (including backpropagation) as well as implementation of these concepts in Python using libraries. Though it spends most of the time looking at supervised learning, unsupervised learning (PCA and $k$-means) as well as some reinforcement learning is also discussed. The course also includes a significant ethics component.

\bigskip

\textbf{Computational Data Science} (\emph{Southern Virginia University CSC 333, upper division computer science course computational data science}) I developed the curriculum to focus on developing practical skills in data science using Python. The course covers data cleaning, visualization, and analysis using libraries such as Pandas, Matplotlib, Seaborn, and some Scikit-learn. The big concept discussed was multiple regression and how to interpret the results of a regression analysis.

\bigskip

\textbf{Mathematical Modeling} (\emph{University of Arizona Math 485, senior level project-based course mathematical modeling})
I assisted Dr. Joceline Lega in converting the course to an asynchronous, fully online format. My principle task was to use Manim (a Python package for programmatic mathematical animation) to create animation to improve the accessibility of the interactive video presentations.

\section*{Mentoring}
\subsection*{Southern Virginia University}
\begin{itemize}
    \item I mentored two undergraduate seniors in their {\bf senior capstone projects}.
    \begin{itemize}
        \item Caden Eskelsen (Spring 2025, project: \emph{Passive and Active Investment:
        A Look At Mutual Fund Performance})
        \item Reilly Kartchner (Fall 2023, project: \emph{The Mathematics of Neural Networks})
    \end{itemize}
  \end{itemize}

\subsection*{Department of Mathematics, University of Arizona}
\begin{itemize}
    \item As part of the \textbf{UA RTG-REU} 2021 summer program and again in summer 2022 (both organized by Dr. Kevin Lin and Dr. Laura Miller), I mentored a total of three undergraduate researchers
    \begin{itemize}
        \item David Parra
        \item Benjamin Shearer
        \item Darianne Sanchez
    \end{itemize}
    \item I mentored three \textbf{undergraduate research groups} as part of a senior level modeling course.
    \begin{itemize}
        \item Spring 2022. Students: Daniel Peck, Aliyah Postell, Marc Bulau, Khaled Ali, Katie Fitts
        \item Spring 2020. Students: Sean Raglow, Gordon Downs, Wes Johnson, Ryan Lewis
        \item Spring 2018. Students: Haoliang Shi, Xiaoyu Shi, Haozhe Xu, Ziwei Yu
    \end{itemize}
    \item As part of the \textbf{Graduate Student Mentor Program} I mentored three students
    \begin{itemize}
        \item John Park
        \item Florence Dungan
        \item Jalen Cates
    \end{itemize}
\end{itemize}

\section*{Service}
\begin{itemize}
    \item Curriculum Committee member at Southern Virginia University (2023 - present).
    \item Graduate representative on the undergraduate committee for the University of Arizona Mathematics Department (2022 - 2023).
\end{itemize}

\section*{Research Interests}
General: Data-driven model reduction, signals and systems theory, stochastic models and stochastic differential equations, power distribution systems with renewable energy sources, smart grids, cell movement, urban traffic.

Current projects: Applying classical signal processing theory to data-driven model reduction, power spectrum estimation, Model reduction for power systems networks, and modeling movement of the slime mold \emph{Dictyostelium discoideum}.

\section*{Publications}
Jared McBride and Kevin K Lin, \emph{A comparison of spectral estimation methods for the analysis of chaotic and stochastic dynamical systems}. In preparation, 2023.

Darin J Law, Jason P Field, Luke A Wilson, Mallory F Barnes, David D Breshears, Jared McBride, Greg A Barron-Gafford, Angelina Martínez-Yrízar, Alberto Búrquez, and Enriquena Bustamante Ortega. \emph{Heatwave compounded warmer drought: Shifts in time to tree mortality and tree mortality duration.} In preparation, 2023.

\section*{Conferences}
\textbf{Third Symposium on Machine Learning and Dynamical Systems}, September 26 - 30, 2022, The Fields Institute, Toronto, Canada\newline
Presented poster: \emph{A comparison of spectral estimation methods for the analysis of chaotic and stochastic dynamical systems}

\section*{Volunteer Community Involvement}
\textbf{Church of Jesus Christ of Latter-day Saints} Full-time Missionary. Manchester, England, 2007-2009

\textbf{Central Utah STEM Fair} Judge. Provo, Utah, 2014

\textbf{ACCESS} (BYU big brother/big sister program) big brother volunteer. Provo, Utah, 2014

\textbf{$\pi$-Day} volunteer. Provo, Utah, 2014-2016

\textbf{S.Y.STEM Coalition Science Day} volunteer. Tucson, Arizona, 2016

\textbf{S.Y.STEM Coalition Summer Robotics Camp} facilitator assistant. Tucson, Arizona, 2019

\textbf{Church of Jesus Christ of Latter-day Saints} recent involvement in local congregation (ward). Tucson, Arizona,
\begin{itemize}
    \item Nursery Leader 2016 - 2017
    \item Sunday School leadership 2017 - 2018
    \item Men's organization (Elders Quorum) leadership 2018
    \item Ward Mission Leader 2018-2019
    \item Assistant Ward Clerk (Membership) 2019 - 2023
    \item Primary Chorister 2023
\end{itemize}

\section*{Skills}
Git, Julia, Python, Manim (Python package for programmatic mathematical animation).

\end{document}